Universal Properties in Generalized Multicategories

### 2018-02-07 upgm.tex
[gen. multicats] some small edits based on recent thoughts

In more detail a representation of $Q : \vcat M \pfrom \cat C$ is a
2-cell:
  \[
  \begin{tikzcd}
TM_0 \arrow[dd, "id"] &  & C \arrow[dd, "G_Q"] \arrow[ll, "Q"] \\
 & I_Q &  \\
TM_0 &  & M_1 \arrow[ll, "M_1"]
  \end{tikzcd}
  \]
that is cartesian in the appropriate sense.
%
Returning to type-theoretic intuition, this square is the
\emph{introduction rule}: above the line we have the ``formal''
product or functin etc, and below the line we have the
``internalized'' notion.
%
The associativity rule gives the definition of substitution for the
introductin rule.
%
We can define an ``elimination'' rule using the cartesian property of
$I_Q$, except \emph{we can't we need a notion of vertical morphism from cat to multicat(!)}:

@@

For left-universal properties, consider sums as an illustrative
example.
%
In a category $\cat C$, a sum/coproduct $A + B$ is defined by the property that
\[ \cat C(A + B, C) \equiv \cat C(A,C) \times \cat C(B,C) \]
%
However, a sum in a multicategory $\vcat C$ should \emph{not} just be
a sum in the underlying category (unlike for right universal
properties).
%
Instead it is what's called a \emph{distributive} coproduct, which can
be defined manually by some laws relating compositions of the
coproduct with the monoidal structure.
%
We want to find an abstract setting that captures these distributivity
properties, which already have a type-theoretic explanation: the
equivalence happens ``in an arbitrary context'':
\[ \vcat C(\Gamma,A + B; C) \equiv \vcat C(\Gamma, A; C) \times C(\Gamma,B;C)\]
this can be defined using fibrations but we will try a different,
hopefully more direct tactic.

>>>>

In more detail a representation of $Q : \vcat M \pfrom \cat C$ is a
2-cell:
  \[
  \begin{tikzcd}
TM_0 \arrow[dd, "id"] &  & C \arrow[dd, "G_Q"] \arrow[ll, "Q"] \\
 & I_Q &  \\
TM_0 &  & M_1 \arrow[ll, "M_1"]
  \end{tikzcd}
  \]
that is cartesian in the appropriate sense and a module
\emph{homomorphism} in that the introduction rule turns $Q$-cuts into
$M$-cuts.
%
Returning to type-theoretic intuition, this square is the
\emph{introduction rule}: above the line we have the ``formal''
product or functin etc, and below the line we have the
``internalized'' notion.
%
The homomorphism property gives the definition of substitution for the
introductin rule.
%
We can define an ``elimination'' rule using the cartesian property of
$I_Q$, except \emph{we can't we need a notion of vertical morphism from cat to multicat(!)}:

@@

For left-universal properties, consider sums as an illustrative
example.
%
In a category $\cat C$, a sum/coproduct $A + B$ is defined by the property that
\[ \cat C(A + B, C) \equiv \cat C(A,C) \times \cat C(B,C) \]
%
However, a sum in a multicategory $\vcat C$ should \emph{not} just be
a sum in the underlying category (unlike for right universal
properties).
%
Instead it is what's called a \emph{distributive} coproduct, which can
be defined manually by some laws relating compositions of the
coproduct with the monoidal structure.
%
We want to find an abstract setting that captures these distributivity
properties, which already have a type-theoretic explanation: the
equivalence happens ``in an arbitrary context'':
\[ \vcat C(\Gamma,A + B; C) \equiv \vcat C(\Gamma, A; C) \times C(\Gamma,B;C)\]
the usual approach is to use fibrations, and our approach is similar,
but more general and we hope more conceptually direct.
